---
title: "World-Happiness Appendix"
author: "Rylan Marianchuk"
date: "STAT 429 project"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(formatR)
library(knitr)
library(ggplot2)
library("car")
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Model Fitting
# Research question 1
This is approached using a 1-Way ANOVA, as described in 1.3.

The hypothesis test follows the null and alternative:
$$H_0:  \mu_{africa} = \mu_{asia} = \mu_{europe} = \mu_{africa} = \mu_{oceania} = \mu_{NAmerica} = \mu_{SAmerica}$$
$$H_A: \mu_i \neq \mu_j \> \texttt{  for some } i,j$$


```{r }
world.happiness <- read.csv("final.csv")

# ANOVA
table(world.happiness$region)

anova = data.frame("Happiness Score" = world.happiness$happiness.score, "Continents" = world.happiness$region)

summary(aov(Happiness.Score ~ Continents, data = anova))

fit <- lm(Happiness.Score ~ Continents, data = anova)
summary(fit)

```

Checking for equality of variances, the ratio of $\frac{\texttt{max}}{\texttt{min}}$ variances of the group happiness distributions, and Bertlett and LeveneTest:

```{r}
# Distributions of each region
africa_happiness <- world.happiness$happiness.score[world.happiness$region == "Africa"]
asia_happiness <- world.happiness$happiness.score[world.happiness$region == "Asia"]
europe_happiness <- world.happiness$happiness.score[world.happiness$region == "Europe"]
namerica_happiness <- world.happiness$happiness.score[world.happiness$region == "North America"]
oceania_happiness <- world.happiness$happiness.score[world.happiness$region == "Oceania"]
samerica_happiness <- world.happiness$happiness.score[world.happiness$region == "South America"]

var_vec = c(var(africa_happiness), var(asia_happiness), var(europe_happiness), var(namerica_happiness), var(oceania_happiness), var(samerica_happiness))

max(var_vec) / min(var_vec)

bartlett.test(Happiness.Score ~ Continents, data = anova)
leveneTest(Happiness.Score ~ Continents, data = anova)
```

See section 2.4 Model Fitting for results discussed in report.


# Research question 2

Performing a multiple linear regression of GDP as a response.

```{r}
mlr.frame <- data.frame("GDP" = world.happiness$GDP, "social" = world.happiness$social, "health" = world.happiness$health,
                        "freedom" = world.happiness$freedom, "trust" = world.happiness$trust, "generosity" = world.happiness$generosity)

full_fit <- lm(GDP ~ social + health + freedom + trust + generosity, data=mlr.frame)
summary(full_fit)

#Arriving at the same coefficients using matrix vector operations to confirm its correctness:
data_mat <- data.matrix(mlr.frame)
p <- 5
n <- nrow(data_mat)
design_mat <- cbind(rep(1,n),data_mat[,c("social", "health", "freedom", "trust", "generosity")])

y.vector <- as.vector(data_mat[,"GDP"])

# Using r matrix operations to solve the beta hat vector
betahat.vector <- solve(t(design_mat) %*% design_mat) %*% t(design_mat) %*% y.vector
betahat.vector

# Finding SSE and residual vector
e.residual <- y.vector - design_mat %*% betahat.vector
SSE <- t(e.residual) %*% e.residual
SSE

s.square <- SSE / (n-p-1)
s <- sqrt(s.square)
s
```


Performing bidirectional variable selection using both AIC and BIC criterion to arrive at 2 candidate models. 

```{r}
# Variable selection and model refinement
  # Using the criterion of AIC
step(full_fit, data = world.happiness, direction = "both",k=2)
  # Using the criterion of BIC
step(full_fit, data = world.happiness, direction = "both",k=log(nrow(world.happiness)))
```

Reduced candidate models are checked by $C_p$ statistic:
```{r}
# 2 candidate models: ------------------

# Model 1 - social + health + trust + generosity
cand1 <- lm(GDP ~ social + health + trust + generosity, data = world.happiness)
summary(cand1)
#Cp Statistic

# Using the matrix operation to get SSE
p <- 4
n <- nrow(data_mat)
design_mat <- cbind(rep(1,n),data_mat[,c("social", "health", "trust", "generosity")])

y.vector <- as.vector(data_mat[,"GDP"])
beta_hat.cand1 <- solve( t(design_mat) %*% design_mat) %*% t(design_mat) %*% y.vector
e.cand1 <- y.vector - design_mat %*% beta_hat.cand1

SSE = t(e.cand1) %*% e.cand1
c(SSE, sum(cand1$residuals^2))
# 0.2346 is the residual standard error of full_fit
Cp1 <- SSE / (0.2346)^2 - (n-2*p)
Cp1


# Model 2 - social + health + trust
cand2 <- lm(GDP ~ social + health + trust, data = world.happiness)
summary(cand2)
# Cp statistic
design_mat <- cbind(rep(1,n),data_mat[,c("social", "health", "trust")])

y.vector <- as.vector(data_mat[,"GDP"])
beta_hat.cand2 <- solve( t(design_mat) %*% design_mat) %*% t(design_mat) %*% y.vector
e.cand2 <- y.vector - design_mat %*% beta_hat.cand2

SSE = t(e.cand2) %*% e.cand2
c(SSE, sum(cand2$residuals^2))
p <- 3
Cp2 <- SSE / (0.2346)^2 - (n-2*p)
Cp2
```

Choice of model using ANOVA:

```{r}

```




